---
title: "analysis"
---

First we load the packages:

```{r, echo = FALSE}
#| output: false
library(babynames)
library(knitr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(pheatmap)

#echo = FALSE : only the output are shown, not the code
#|# output: false : the output are not shown
```

The table below display the first 10 rows of babynames:

```{r}
head(babynames) |> kable()

```

To visualize the most popular baby names, two functions were created (get_most_frequent() and plot_top())

```{r}
#| code-fold: true
get_most_frequent <- function(babynames, select_sex, from = 1950) {
  most_freq <- babynames |>
    filter(sex == select_sex, year > from) |>
    group_by(name) |>
    summarise(average = mean(prop)) |>
    arrange(desc(average))
    
  return(list(
    babynames = babynames,
    most_frequent = most_freq,
    sex = select_sex,
    from = from))
}

plot_top <- function(x, top = 10) {
  topx <- x$most_frequent$name[1:top]
  
  p <- x$babynames |>
    filter(name %in% topx, sex == x$sex, year > x$from) |>
    ggplot(aes(x = year, y = prop, color = name)) +
    geom_line() +
    scale_color_brewer(palette = "Paired") +
    theme_classic()
  
  return(p)
}
```

We used these functions to visualize the 10 most frequent girl names from 1950 onwards:

```{r}
#| label: fig-line-girls
#| fig-cap: "Line plot displaying proportion of top 10 girl names by year"
get_most_frequent(babynames, select_sex = "F") |>
  plot_top()
```

In @fig-line-girls you can see that there has been a peak in popularity for 'Lisa', 'Jennifer' and 'Jessica'. Interesting! Let's have a look at the boys names:

```{r}
#| label: fig-line-boys
#| fig-cap: "Line plot displaying proportion of top 10 boy names by year"
get_most_frequent(babynames, select_sex = "M") |>
  plot_top()
```

## @fig-line-boys shows that names that were popular before 1990 are relatively infrequent after 2000.
